# Reproducible Research: Peer Assessment 1
========================================================

## Loading and preprocessing the data
```{r echo=TRUE}
data<-read.csv("activity.csv",sep=",",stringsAsFactor=F,header=TRUE)
data$date<-as.Date(data$date,format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
library(sqldf)
data1<-data[which(data$steps!="NA"),]
data1sum<-sqldf('select data1.date, sum(data1.steps) as totsteps from data1 group by data1.date')
hist(data1sum$totsteps, col="gray", main="Histogram of the total number of steps taken each day", ylab="", xlab="total number of steps taken each day", breaks = 9)
x<-mean(data1sum$totsteps)
y<-median(data1sum$totsteps)
```
The mean total number of steps taken per day is `r x`.
The median total number of steps taken per day is `r y`.

## What is the average daily activity pattern?
```{r echo=TRUE}
data2ave<-sqldf('select data1.interval, avg(data1.steps) as avgsteps from data1 group by data1.interval')
plot(data2ave$interval,data2ave$avgsteps, type="l", ylab="average number of steps taken", xlab="5-minute interval", main="Average daily activity pattern")
data2max<-data2ave[order(-data2ave$avgsteps),]
z<-data2max[1,2]
```
`r z` is 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.

## Imputing missing values
```{r echo=TRUE}
n<-nrow(data)-nrow(data1)
```

The total number of missing values in the dataset is `r n`.  
The strategy for filling in all of the missing values in the dataset is the mean for that 5-minute interval.  
Create a new dataset that is equal to the original dataset but with the missing data filled in:
``` {r echo=TRUE}
data3<-sqldf('select data.*, data2ave.avgsteps from data left outer join data2ave on data.interval=data2ave.interval')
data3$stepcln<-ifelse(data3$steps %in% (NA), data3$avgsteps, data3$steps)
datacln<-data3[,c("date","interval","stepcln")]
data2sum<-sqldf('select datacln.date, sum(datacln.stepcln) as totsteps from datacln group by datacln.date')
par(mfrow=c(2,1))
hist(data1sum$totsteps, col="gray", main="Histogram of the total number of steps taken each day (NA ignored)", ylab="", xlab="total number of steps taken each day", breaks = 9)
hist(data2sum$totsteps, col="gray", main="Histogram of the total number of steps taken each day (NA cleaned)", ylab="", xlab="total number of steps taken each day", breaks = 9)
x2<-mean(data2sum$totsteps)
y2<-median(data2sum$totsteps)
```
The mean total number of steps taken per day is `r x2`.  
The median total number of steps taken per day is `r y2`.

## Are there differences in activity patterns between weekdays and weekends?

``` {r echo=TRUE}
datacln$weekd<-weekdays(datacln$date)
datacln$weekd<-ifelse(datacln$weekd %in% ("суббота") | datacln$weekd %in% ("воскресенье"), "weekend","weekday")
datawd<-sqldf('select datacln.weekd, datacln.interval, avg(datacln.stepcln) as number_of_steps from datacln group by datacln.weekd, datacln.interval')
datawd$weekd<-as.factor(datawd$weekd)
library(lattice)
p<-xyplot(number_of_steps~interval|weekd,datawd, layout=c(1,2),type="l")
print(p)

```
